Confirmation of Wide-Field Signatures in Redshifted 21~cm Power Spectra

### 2015-06-05 ms.tex
Modify sentence on power spectrum estimation
From eea4f3e:
============

In ms.tex,

(*) Modify sentence on optimum power spectrum estimation in Section 6
Minor changes throughout
From 29fc971:
============

In ms.tex,

(*) Minor changes in Introduction
(*) Minor changes to Section 3
(*) Minor changes to Section 5

A careful design of antenna aperture can significantly mitigate this contamination. Optimal weighting of foreground-contaminated Fourier modes may be required to extract the signal with maximum sensitivity. Thus, knowledge of such detailed foreground signatures is key for design and analysis choices of future instruments such as HERA and SKA.

@@

To reduce thermal fluctuations while maintaining coherence, it is essential to average independent data sets obtained over the same region of sky with identical instrument settings. Hence, we select a subset of MWA snapshots each of duration 112~seconds obtained over different nights which are aligned to within 72~seconds of each other in {\it local sidereal time} (LST) around a mean LST of 0.04~hours with the MWA tile beam pointed at zenith. The database contains 14 snapshots satisfying these criteria. Two of these snapshots, which contained amplitude and phase artifacts for a significant duration across different baselines, are excluded from our analysis. Results of this coherent averaging are discussed below.

@@

We approached this issue in Paper~I via simulations of different antenna apertures, namely, a dipole, a $4\times 4$ phased array, and a dish. Among these, a dish provides the best quality as evaluated from a foreground contamination viewpoint. With new instruments such as HERA and SKA on the horizon, we extend this discussion to further emphasize the need to comprehensively study the effect of their antenna apertures on foreground contamination.

@@

With the SKA at low frequencies, deploying ``stations'' of $\sim$~35~m diameter that act as aperture arrays consisting of 256 pseudo-randomly placed vertical log-periodic dipole antennas is being considered\footnote{Documents at \url{https://www.skatelescope.org/key-documents/} on SKA re-baselining}$^,$\footnote{SKA memo ``Station Response and Imaging Performance of LFAA: 100 MHz to 600 MHz'' by Razavi-Ghods et al.}. The voltage beam of the station will be obtained by a phased addition of the dipole responses. Since the power pattern of a baseline consisting of two stations will be effectively a product of two voltage beams, it will tend to homogenize the fluctuations in the power patterns of individual stations. Even with simple models of such phased array stations, we estimate the typical horizon response of these beams to be $-30$~dB to $-40$~dB compared to zenith, yielding a response that is at least 10~dB better than the MWA. Targeted optimization of antenna layouts of stations is under active study and could yield even better responses.

From these simple antenna aperture models, HERA and SKA should mitigate leakage from the {\it pitchfork} into the {\it EoR window} by $\sim 40$~dB relative to the MWA when expressed in units of delay power spectrum. This will significantly increase the likelihood of detecting the cosmological H~{\sc i} power spectrum in sensitive three-dimensional $k$-modes adjacent to the {\it pitchfork}. The precise beam responses and the resulting foreground leakage, especially from near the horizon, will prove to be critical and timely inputs to the actual aperture designs of HERA dishes and SKA stations which are nearing their final stages.

@@

Foreground and instrument models serve as inputs for power spectrum estimation techniques \citep[for e.g.,][]{liu14a,liu14b,dil15}. Input models which ignore wide-field effects will provide over-valued weights in and around the {\it pitchfork} modes, and hence result in sub-optimal results. Thus, confirmation of the {\it pitchfork} effect has important implications for power spectrum estimation.

>>>>

Careful antenna aperture design can significantly mitigate this contamination. Optimal weighting of contaminated Fourier modes may be required to extract the signal with maximum sensitivity. Thus, detailed knowledge of foreground signatures is key for design and analysis choices of future instruments such as HERA and SKA.

@@

To reduce thermal fluctuations while maintaining coherence, it is essential to average independent data sets obtained over the same region of sky with identical instrument settings. Hence, we select a subset of MWA snapshots pointed at zenith each of duration 112~seconds obtained over different nights which are aligned to within 72~seconds of each other in {\it local sidereal time} (LST) around a mean LST of 0.04~hours. The database contains 14 snapshots satisfying these criteria. Two of these snapshots, which contained amplitude and phase artifacts for a significant duration across different baselines, are excluded from our analysis. Results of this coherent averaging are discussed below.

@@

We approached this issue in Paper~I via simulations of different antenna apertures -- a dipole, a $4\times 4$ phased array, and a dish. Among these, a dish provides the best quality as evaluated from a foreground contamination viewpoint. With new instruments such as HERA and SKA on the horizon, we extend this discussion to further emphasize the need to comprehensively study the effect of their antenna apertures on foreground contamination.

@@

With the SKA at low frequencies, deploying ``stations'' of $\sim$~35~m diameter that act as aperture arrays consisting of 256 pseudo-randomly placed vertical log-periodic dipole antennas is being considered\footnote{Documents at \url{https://www.skatelescope.org/key-documents/} on SKA re-baselining}$^,$\footnote{SKA memo ``Station Response and Imaging Performance of LFAA: 100 MHz to 600 MHz'' by Razavi-Ghods et al.}. The voltage beam of the station will be obtained by a phased addition of the dipole responses. Since the power pattern of a baseline consisting of two stations will be effectively a product of two voltage beams, it will homogenize the fluctuations in the power patterns of individual stations. Even with simple models of such phased array stations, we estimate the typical horizon response of these beams to be $-30$~dB to $-40$~dB compared to zenith, yielding a response that is at least 10~dB better than the MWA. Targeted optimization of antenna layouts of stations is under active study and could yield even better responses.

From these simple antenna aperture models, HERA and SKA should mitigate leakage from the {\it pitchfork} into the {\it EoR window} by $\sim 40$~dB relative to the MWA when expressed in units of power spectrum. This will significantly increase the likelihood of detecting the cosmological H~{\sc i} power spectrum in sensitive three-dimensional $k$-modes adjacent to the {\it pitchfork}. The precise beam responses and the resulting foreground leakage, especially from near the horizon, will prove to be critical and timely inputs to the actual aperture designs of HERA dishes and SKA stations which are nearing their final stages.

@@

Foreground and instrument models serve as inputs for power spectrum estimation techniques \citep[for e.g.,][]{liu14a,liu14b,dil15}. Input models which ignore wide-field effects will provide over-valued weights in and around the {\it pitchfork} modes, and hence result in sub-optimal results. Thus, confirmation of the {\it pitchfork} effect has important implications for optimum power spectrum estimation.
